The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2016-02-14 script/en/chapters/ellipticcurves.tex
CT Script: Minor changes (typo)
First version of CT Book ed 12. Most important point is the newly includes bitciphers chapter. Further things: Tried to reduce latex warnings amap. Changed wording from Sage to SageMath where visible...

This section provides information about \index{Group} {\em groups} and
\index{Field} {\em fields}.

@@

Fields may contain an infinite number of elements (e.g.\ the field of real numbers). They are called {\em infinite} fields. In contrast we call a field
{\em finite}, if it contains only a finite number of elements (e.g.\ ${\mathbb Z}_p = \{0, 1, 2, 3, \cdots, p-1\}$
, where $p$ is a prime. ${\mathbb Z}_p$ with addition mod $p$ and multiplication
mod $p$).
\index{Field!characteristic}
\paragraph*{Characteristic of a field}
Let $K$ be a field and $1$ be the neutral element of $K$ with
respect to the multiplicative operation ``$\cdot$''. Then the characteristic of $K$ is said to be the order of $1$ with respect to the additive operation. This means that the characteristic of $K$ is the smallest positive integer $n$ such that
$$ \underbrace{1+1+\dots+1}_{\hbox{$n$ times}} =0 .
$$
If there is no such $n$, i.e. if $1+1+\dots+1\ne 0$ no matter how many $1$s we add, then we call $K$ a field
with characteristic $0$.

@@

\index{Field!finite}
\paragraph*{Finite Fields}
As mentioned above, each finite field has a characteristic $p\ne 0$, where $p$ is a prime. On the other hand, given a prime $p$ there is a field which has exactly $p$ elements, that is ${\mathbb Z}_p$.

@@

There are not many free programms offering ECC under a graphical user interface.
The following subsections explain which according functionality is available in
CrypTool and in Sage.

@@

% -----------------------------------------------------------------------------
\subsection{Sage}
\label{ec:Sage_Massierer}
\index{Sage}
\index{Sage!Code examples}

In Sage elliptic curves are described at

\url{http://www.sagemath.org/doc/constructions/elliptic_curves.html}%
\footnote{%
According Sage samples can be found at the "Published Worksheets" at
\url{http://www.sagenb.org/pub/}\\
- about Elliptic Curve: \url{http://www.sagenb.org/home/pub/606/} \\
- about Elliptic Curve El Gamal: \url{http://www.sagenb.org/home/pub/104/}, or at

@@

\noindent Additionally there is an exhaustive, interactive
\hyperlink{ec:Web-Link:Sage_Massierer}{ECC tutorial} by Maike Massierer.
This interactive introduction to Elliptic Curve Cryptography is built up as
a Sage notebook.

Sage notebooks are running after a logon within a browser%
\footnote{%
If you installed Sage on your own (Unix) server, you first have to enter at the
command line the command \verb#notebook()#.

}${}^,$\footnote{%
The ECC notebook of Maike Massierer needs the KASH3 library: Therefore e.g.\ with Sage 4.2.1
the package ``kash3-2008-07-31.spkg'' has to be installed (command \verb#sage -i#).
}.

The ECC notebook\index{Elliptic curves!ECC notebook}\index{Massierer, Maike} of
Massierer\footnote{%
Instructions to use an interactive Sage notebook\index{Sage!Instructions interactive notebook}:\\

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Public Sage servers like \url{http://sage.mathematik.uni-siegen.de:8000} or
  \url{http://www.sagenb.org/} often offer running samples as "Published Worksheets",
  which you can run and download without login on. These worksheets are listed
  if you click on ``Published'' in the above right corner.

@@

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Sign-up for a Sage notebook account at \url{http://sagenb.org/register}
  and sign in at \url{http://sagenb.org/}.

@@

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  According to the Sage-support newsgroup there is work underway
  to create a system for using \verb#@interact# completely outside of the
  Sage notebook (JS code within a static html pages).

@@

   \hypertarget{ec:Web-Link:Sage_Massierer}{} 
   \item Interactive introduction to elliptic curves and elliptic curve
        cryptography with Sage\index{Sage} 
        by Maike Massierer and the CrypTool team,\\
        \url{http://sagenb.org/home/pub/1126/} (\verb!#1126 to #1133!) \\
        ECC Tutorial as Sage Notebook \\
        Version 1.2, November 2009

>>>>

This section provides information about \index{Group} {\em groups} and
\index{field} {\em fields}.

@@

Fields may contain an infinite number of elements (e.g.\ the field of real numbers). They are called {\em infinite} fields. In contrast we call a field
{\em finite}, if it contains only a finite number of elements (e.g.\ ${\mathbb Z}_p = \{0, 1, 2, 3, \cdots, p-1\}$
, where $p$ is a prime. ${\mathbb Z}_p$ with addition mod $p$ and multiplication
mod $p$).
\index{field!characteristic}
\paragraph*{Characteristic of a field}
Let $K$ be a field and $1$ be the neutral element of $K$ with
respect to the multiplicative operation ``$\cdot$''. Then the characteristic of $K$ is said to be the order of $1$ with respect to the additive operation. This means that the characteristic of $K$ is the smallest positive integer $n$ such that
$$ \underbrace{1+1+\dots+1}_{\hbox{$n$ times}} =0 .
$$
If there is no such $n$, i.e. if $1+1+\dots+1\ne 0$ no matter how many $1$s we add, then we call $K$ a field
with characteristic $0$.

@@

\index{field!finite}
\paragraph*{Finite Fields}
As mentioned above, each finite field has a characteristic $p\ne 0$, where $p$ is a prime. On the other hand, given a prime $p$ there is a field which has exactly $p$ elements, that is ${\mathbb Z}_p$.

@@

There are not many free programms offering ECC under a graphical user interface.
The following subsections explain which according functionality is available in
CrypTool and in SageMath.

@@

% -----------------------------------------------------------------------------
\subsection{SageMath}
\label{ec:Sage_Massierer}
\index{SageMath}
\index{SageMath!Code examples}

In SageMath elliptic curves are described at

\url{http://www.sagemath.org/doc/constructions/elliptic_curves.html}%
\footnote{%
According SageMath samples can be found at the "Published Worksheets" at
\url{http://www.sagenb.org/pub/}\\
- about Elliptic Curve: \url{http://www.sagenb.org/home/pub/606/} \\
- about Elliptic Curve El Gamal: \url{http://www.sagenb.org/home/pub/104/}, or at

@@

\noindent Additionally there is an exhaustive, interactive
\hyperlink{ec:Web-Link:Sage_Massierer}{ECC tutorial} by Maike Massierer.
This interactive introduction to Elliptic Curve Cryptography is built up as
a SageMath notebook.

SageMath notebooks are running after a logon within a browser%
\footnote{%
If you installed SageMath on your own (Unix) server, you first have to enter at the
command line the command \verb#notebook()#.

}${}^,$\footnote{%
The ECC notebook of Maike Massierer needs the KASH3 library: Therefore e.g.\ with SageMath 4.2.1
the package ``kash3-2008-07-31.spkg'' has to be installed (command \verb#sage -i#).
}.

The ECC notebook\index{Elliptic curves!ECC notebook}\index{Massierer, Maike} of
Massierer\footnote{%
Instructions to use an interactive SageMath notebook\index{SageMath!Instructions interactive notebook}:\\

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Public SageMath servers like \url{http://sage.mathematik.uni-siegen.de:8000} or
  \url{http://www.sagenb.org/} often offer running samples as "Published Worksheets",
  which you can run and download without login on. These worksheets are listed
  if you click on ``Published'' in the above right corner.

@@

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  Sign-up for a SageMath notebook account at \url{http://sagenb.org/register}
  and sign in at \url{http://sagenb.org/}.

@@

  \noindent\hangindent=6pt\makebox[6pt][l]{-}%
  According to the SageMath-support newsgroup there is work underway
  to create a system for using \verb#@interact# completely outside of the
  SageMath notebook (JS code within a static html pages).

@@

   \hypertarget{ec:Web-Link:Sage_Massierer}{} 
   \item Interactive introduction to elliptic curves and elliptic curve
        cryptography with Sage\index{SageMath} 
        by Maike Massierer and the CrypTool team,\\
        \url{http://sagenb.org/home/pub/1126/} (\verb!#1126 to #1133!) \\
        ECC Tutorial as SageMath Notebook \\
        Version 1.2, November 2009

### 2016-02-14 script/en/chapters/homomorphicencryption.tex
First version of CT Book ed 12. Most important point is the newly includes bitciphers chapter. Further things: Tried to reduce latex warnings amap. Changed wording from Sage to SageMath where visible...

$Y$ corresponds to the set of cipher texts, $X$ is the set of plaintexts. For two plaintexts $x_1, x_2$ with corresponding ciphertexts $y_1, y_2$ it holds:
%
$$f(y_1  * y_2) = f(y_1) + f(y_2) = x_1  + x_2$$
%
This equation can be interpreted as follows: If we multiply two ciphertexts $y_1, y_2$ with each other and subsequently decrypt their product, then we will obtain the sum of the originally encrypted values $x_1$ and $x_2$. Everybody can -- without knowledge of the plaintexts, without having to decrypt and even without knowing the private decryption key – compute a product of the two ciphertexts and knows that upon decryption the owner of the private key will obtain the sum of the two originally encrypted plaintexts.

>>>>

$Y$ corresponds to the set of cipher texts, $X$ is the set of plaintexts. For two plaintexts $x_1, x_2$ with corresponding ciphertexts $y_1, y_2$ it holds:
%
$$f(y_1  * y_2) = f(y_1) + f(y_2) = x_1  + x_2$$
%
This equation can be interpreted as follows: If we multiply two ciphertexts $y_1, y_2$ with each other and subsequently decrypt their product, then we will obtain the sum of the originally encrypted values $x_1$ and $x_2$. Everybody can -- without knowledge of the plaintexts, without having to decrypt and even without knowing the private decryption key -- compute a product of the two ciphertexts and knows that upon decryption the owner of the private key will obtain the sum of the two originally encrypted plaintexts.

### 2016-02-14 script/en/chapters/introduction.tex
First version of CT Book ed 12. Most important point is the newly includes bitciphers chapter. Further things: Tried to reduce latex warnings amap. Changed wording from Sage to SageMath where visible...

% $Id$
% ............................................................................
%      V O R W O R T  und  E I N F Ü H R U N G (Zusammenspiel Skript-CT) 
%      P R E F A C E  and  I N T R O D U C T I O N (Playing together of skript and CT) 
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% --------------------------------------------------------------------------
\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Preface to the 11th Edition of the CrypTool Script}
\chapter*{Preface to the 11th Edition of the CrypTool Script}

The script's goal is to explain some mathematical topics in more detail, but
still in a way which is easy to understand.

Starting in the year 2000 this script became part of the CrypTool 1
(CT1)\index{CrypTool 1} package in version 1.2.01. 
Since then it has been expanded and revised in almost every new version of CT1.

@@

The later editorial work in TeX added footnotes and cross linkages between
different sections.

%This edition of the CT script again updated some topics:
Compared to edition 10, this edition completely updated the TeX sources of the
document, and of course, the content of the script was amended, corrected, and
updated with some topics, e.g.:
\vspace{-5pt}
\begin{itemize}
  \item the definitions of the strength of security functions
        (chapter \ref{cm_Section_Security_Definitions}),
%  \item the largest prime numbers (chap. \ref{search_for_very_big_primes}),
%        new factorization records (chap. \ref{RSA-768}),
%  \item progress in cryptanalysis of AES 
%        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played a major role (see appendix \ref{s:appendix-movies}),
%       and where primes are used as hangers  
%        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}),
  \item the overviews of all functions in
        \hyperlink{appendix-template-overview-CT2}{CrypTool 2 (CT2)},
        \hyperlink{appendix-function-overview-JCT}{JCrypTool (JCT)}, and
        \hyperlink{appendix-function-overview-CTO}{CrypTool-Online (CTO)}
        (see appendix),
  \item further Sage scripts for cryptography, and the appendix
        \ref{s:appendix-using-sage} about using the computer algebra system
        Sage,
        % Nguyen and Massierer (see also chap.~\ref{ec:Sage_Massierer}).
  \item the section about the Goldbach conjecture
        (see \ref{L-GoldbachConjecture}) and about twin primes
        (see \ref{L-TwinCousinPrimes}),
  \item the section about shared primes in RSA modules used in reality
        (see \ref{nt_Shared-Primes}),
  \item the section about RSA fixed points
        (see \ref{l:NumberTheory_Sage_Number-of-RSA-FixedPoints}), and
  \item homomorphic encryption
        (see chapter \ref{Chapter_HomomorphicCiphers}).
\end{itemize}

@@

Thanks also to the readers who sent us feedback.
% Especially I would like to acknowledge the English language proof-reading
% of this script version done by Richard Christensen and Lowell Montgomery.

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\par \vskip + 15pt

\noindent Bernhard Esslinger
\\
\\
Frankfurt and Siegen (Germany), August 2013
\par \vskip + 30pt

@@

% --------------------------------------------------------------------------
\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Introduction -- How do the Script and the Program Play together?}
\chapter*{Introduction -- How do the Script and the Program Play together?}

\textbf{This CrypTool script}

@@

The articles in this script are largely self-contained and
can also be read independently of the CrypTool programs.

@@

CrypTool 1 (CT1) is an educational program with a comprehensive
online help enabling you to use and analyse cryptographic procedures within a
unified graphical user interface. The online help in CrypTool 1 contains both
instructions how to use the program and explanations about the methods itself
(both not as detailled and in another structure than the CT script).

@@

MTC3 is the abbreviation for MysteryTwister C3, an international cryptography
contest (\url{http://www.mysterytwisterc3.org}), which is also based on the
CT project.
Here you can find cryptographic riddles in four categories, a
high-score list and a moderated forum. In the meantime more than 3500 active
users participate, and more than 140 challenges are offered (currently 89 of
them are solved by at least one user).
\par \vskip + 25pt  % more than +15 to start Sag at the next page

\noindent \textbf{The Computer Algebra System Sage\index{Sage}}

Sage is a comprehensive open-source CAS package which can be used to easily
program the mathematical methods explained in this script. A speciality of
this CAS is, that the scripting language is Python.\\
Sage becomes more and more the standard open-source CAS system at universities.
\par \vskip + 15pt

@@

Within this initiative, one and two 2 day courses in cryptology are offered
for pupils and teachers in order to show how attractive MINT subjects like
mathematics, computer science and especially cryptology are.
The course agenda is a virtual secret agent story.\\
In the meantime, these courses took place for several years in Germany in
different towns.\\
All course material is freely available on net
(\url{http://www.cryptool.org/schuelerkrypto/}).\\
All software used is free software (using mostly CT1 and CT2).\\
As all course material is currently only available in German we'd be happy
if someone could translate the course material and build an according course
in English.
\par \vskip + 15pt

@@

\noindent Bernhard Esslinger
\\
\\
Frankfurt and Siegen (Germany), August 2013

>>>>

% $Id$
% ............................................................................
%      V O R W O R T  und  E I N F U E H R U N G (Zusammenspiel CT-Buch und CT-Programmen) 
%      P R E F A C E  and  I N T R O D U C T I O N (Playing together of book and programs) 
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% --------------------------------------------------------------------------
\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Preface to the 12th Edition of the CrypTool Book xxxxxxxxxxxx}
\chapter*{Preface to the 12th Edition of the CrypTool Book xxxxxxxxxxxx}

The book's goal is to explain some mathematical topics from cryptology in
exact detail, nevertheless in a way which is easy to understand.

Starting in the year 2000 this book became part of the CrypTool 1
(CT1)\index{CrypTool 1} package in version 1.2.01. 
Since then it has been expanded and revised in almost every new version of CT1.

@@

The later editorial work in LaTeX added footnotes and cross linkages between
different sections, and some corrections.

%This edition of the CT book again updated some topics:
Compared to edition 11, in this edition
% completely updated the TeX sources of the document, and of course,
the content of the book was amended, corrected, and
updated with some topics, for instance:
\vspace{-5pt}
\begin{itemize}
%  \item the definitions of the strength of security functions
%        (chapter \ref{cm_Section_Security_Definitions}),
  \item the largest prime numbers (chap. \ref{search_for_very_big_primes}),
%        new factorization records (chap. \ref{RSA-768}),
%  \item progress in cryptanalysis of AES 
%        (chap. \ref{NeueAES-Analyse}) and
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played a major role (see appendix \ref{s:appendix-movies}),
%       and where primes are used as hangers  
%        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}),
  \item the overviews of all functions in
        \hyperlink{appendix-template-overview-CT2}{CrypTool 2 (CT2)},
        \hyperlink{appendix-function-overview-JCT}{JCrypTool (JCT)}, and
        \hyperlink{appendix-function-overview-CTO}{CrypTool-Online (CTO)}
        (see appendix),
  \item further SageMath scripts for cryptography, and the appendix
        \ref{s:appendix-using-sage} about using the computer algebra system
        SageMath,
        % Nguyen and Massierer (see also chap.~\ref{ec:Sage_Massierer}).
  \item the section about the Goldbach conjecture
        (see \ref{L-GoldbachConjecture}) and about twin primes
        (see \ref{L-TwinCousinPrimes}),
  \item the section about shared primes in RSA modules used in reality
        (see \ref{nt_Shared-Primes}),
%  \item the section about RSA fixed points
%        (see \ref{l:NumberTheory_Sage_Number-of-RSA-FixedPoints}), and
%  \item homomorphic encryption
%        (see chapter \ref{Chapter_HomomorphicCiphers}).
  \item the \nameref{Chapter_BitCiphers} is completely new
        (see chapter \ref{Chapter_BitCiphers}).
\end{itemize}

@@

Thanks also to the readers who sent us feedback.
% Especially I would like to acknowledge the English language proof-reading
% of this book version done by Richard Christensen and Lowell Montgomery.

I hope that many readers have fun with this book and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\par \vskip + 15pt

\noindent Bernhard Esslinger
\\
\\
Frankfurt and Siegen (Germany), February 2016
\par \vskip + 30pt

@@

% --------------------------------------------------------------------------
\clearpage\phantomsection
\addcontentsline{toc}{chapter}{Introduction -- How do the Book and the Programs Play together?}
\chapter*{Introduction -- How do the Book and the Programs Play together?}

\textbf{This CrypTool book}

@@

The chapters in this book are largely self-contained and
can also be read independently of the CrypTool programs.

@@

CrypTool 1 (CT1) is an educational program with a comprehensive
online help enabling you to use and analyse cryptographic procedures within a
unified graphical user interface. The online help in CrypTool 1 contains both
instructions how to use the program and explanations about the methods itself
(both not as detailled and in another structure than the CT book).

@@

MTC3 is the abbreviation for MysteryTwister C3, an international cryptography
contest (\url{http://www.mysterytwisterc3.org}), which is also based on the
CT project.
Here you can find cryptographic riddles in four categories, a high-score list
and a moderated forum. In the meantime more than 4000 active users participate,
and more than 150 challenges are offered (currently 109 of them are solved by
at least one user).xxxxxxxxxxxxxx
\par \vskip + 25pt  % more than +15 to start Sag at the next page

\noindent \textbf{The Computer Algebra System SageMath\index{SageMath}}

SageMath is a comprehensive open-source CAS package which can be used to easily
program the mathematical methods explained in this book. A speciality of
this CAS is, that the scripting language is Python.\\
SageMath becomes more and more the standard open-source CAS system at universities.
\par \vskip + 15pt

@@

Within this initiative, one and two 2 day courses in cryptology are offered
for pupils and teachers in order to show how attractive MINT subjects like
mathematics, computer science and especially cryptology are.
The course agenda is a virtual secret agent story.\\
In the meantime, these courses took place for several years in Germany in
different towns.\\
All course material is freely available on net
(\url{http://www.cryptool.org/schuelerkrypto/}).\\
All software used is free software (using mostly CT1 and CT2).\\
As all course material is currently available only in German -- we'd be happy
if someone could translate the course material and build an according course
in English.
\par \vskip + 15pt

@@

\noindent Bernhard Esslinger
\\
\\
Frankfurt/Siegen (Germany), February 2016

### 2016-02-14 script/en/chapters/moderncryptography.tex
CT Script: Added 48th M-Prime and some minor changes (DE and EN)
First version of CT Book ed 12. Most important point is the newly includes bitciphers chapter. Further things: Tried to reduce latex warnings amap. Changed wording from Sage to SageMath where visible...

\begin{cryptoprocedure}
\textbf{Key generation:} \vskip + 5pt

Let $ p $ and $ q $ be two different prime numbers and $ N = pq$. Let 
$ e $ be any prime number relative to $ \phi (N) $
\index{Prime number!relative prime}\index{Relative prime},
i.e. $ \ggt (e,\phi (N))=1. $\index{gcd}
Using the Euclidean algorithm, 
we calculate the natural number  $ d < \phi (N), $ such that
\[ ed \equiv 1 \mod \phi (N). \]
whereby $ \phi $ is the {\bf Euler phi Function}.

@@

\textbf{Public key:}
\[ N,e. \]
\textbf{Private key:}
\[ d. \]
\textbf{Encryption:}
\[ y= e_{T} (x) = x^{e} \mod N.\]
\textbf{Decryption:}
\[ d_{T} (y) = y^d \mod N. \]
\caption{RSA (based on the factorization problem)}
\label{rsaproc}
\index{Factorization!factorization problem}
\index{Euler!(phi) function}
\end{cryptoprocedure}

@@

\vskip +10 pt
\begin{Proof}{}
Let $ k=2^n + k_{n-1} 2^{n-1} + \cdots + k_1 2 + k_0 $ be the binary 
representation of $k. $ Then $ n \leq \log_2 (k), $  because $ 2^n \leq k < 
2^{n+1}. $ $ k $ can be written in the form $ k=2k' + k_0 $ with $ k'= 2^{n-1} + 
k_{n-1} 2^{n-2} + \cdots + k_1 $. Thus 
\[ g^k = g^{2k'+k_0}= (g^{k'})^2 g^{k_0} .\]
We therefore obtain $ g^k $ from $ g^{k'} $ by squaring and then multiplying by 
$ g $. The claim is thus proved by induction to $ n. $
\end{Proof}

@@

\begin{Proof}{}
  First create the two lists \\
Giant-step list: $ \{1,g^n,g^{2n}, \ldots, g^{n \cdot n}\}, $\\
Baby-step list: $ \{ hg^{-1} , hg^{-2} , \ldots , hg^{-n} \}. $ \par
If $ g^{jn} = hg^{-i}, $ i.e. $ h = g^{i+jn}, $ then the problem is solved. If 
the lists are disjoint, then $ h $ cannot be represented as $ g^{i + jn}, i, 
j\leq n,$. As all powers of $ g $ are thus recorded, the logarithm problem does 
not have a solution.
\end{Proof}

>>>>

\index{Factorization!factorization problem}
\index{Euler!(phi) function}
\begin{cryptoprocedure}
\textbf{Key generation:} \vskip + 5pt

Let $ p $ and $ q $ be two different prime numbers and $ N = pq$.
Let $ e $ be any number relative prim to $ \phi (N) $
\index{Prime number!relative prime}\index{Number!relative prime},
i.e. $ \ggt (e,\phi (N))=1. $\index{gcd}
Using the Euclidean algorithm, 
we calculate the natural number  $ d < \phi (N), $ such that
\[ ed \equiv 1 \mod \phi (N). \]
whereby $ \phi $ is the {\bf Euler phi Function}.

@@

\textbf{Public key:}
\[ N,e. \]
\textbf{Private key:}
\[ d. \]
\textbf{Encryption:}
\[ y= e_{T} (x) = x^{e} \mod N.\]
\textbf{Decryption:}
\[ d_{T} (y) = y^d \mod N. \]
\caption{RSA (based on the factorization problem)}
\label{rsaproc}
\end{cryptoprocedure}

@@

\vskip +10 pt
\begin{proof}{}
Let $ k=2^n + k_{n-1} 2^{n-1} + \cdots + k_1 2 + k_0 $ be the binary 
representation of $k. $ Then $ n \leq \log_2 (k), $  because $ 2^n \leq k < 
2^{n+1}. $ $ k $ can be written in the form $ k=2k' + k_0 $ with $ k'= 2^{n-1} + 
k_{n-1} 2^{n-2} + \cdots + k_1 $. Thus 
\[ g^k = g^{2k'+k_0}= (g^{k'})^2 g^{k_0} .\]
We therefore obtain $ g^k $ from $ g^{k'} $ by squaring and then multiplying by 
$ g $. The claim is thus proved by induction to $ n. $
\end{proof}

@@

\begin{proof}{} !N!*
  First create the two lists \\
Giant-step list: $ \{1,g^n,g^{2n}, \ldots, g^{n \cdot n}\}, $\\
Baby-step list: $ \{ hg^{-1} , hg^{-2} , \ldots , hg^{-n} \}. $ \par
If $ g^{jn} = hg^{-i}, $ i.e. $ h = g^{i+jn}, $ then the problem is solved. If 
the lists are disjoint, then $ h $ cannot be represented as $ g^{i + jn}, i, 
j\leq n,$. As all powers of $ g $ are thus recorded, the logarithm problem does 
not have a solution.
\end{proof}
